### 2016-06-23 motivation.tex
Edits.
edits

\begin{figure}[!h]
\centering
\subcaptionbox {
  {\sc rc} Execution 1
  \label{fig:motiv-eg-1-a}
} [
  0.55\columnwidth
] {
  \includegraphics[scale=0.5]{Figures/motiv-eg-1-a}
}
%\hspace*{0.5in}
\subcaptionbox {
  {\sc rc} Execution 2
  \label{fig:motiv-eg-1-b}
}{
  \includegraphics[scale=0.5]{Figures/motiv-eg-1-b}
}
\caption{A possible execution of the program shown in Fig.~\ref{fig:motiv-eg-1} under
  a read committed isolation level.}
\end{figure}
The figure depicts an execution as a series of read, write, and commit
operations.  The subscript of an operation indicates the transaction
that executes it\footnote{For clarity, the effects of different
  transactions are shown in different colored backgrounds.} In the
first execution, transaction {\bf Wd1} reads the current balance
(\C{k}) and writes the new balance (\C{k-a1}), but before it commits
transaction {\bf Wd2} executes and commits, writing the new balance
(\C{k-a2}). RC isolation prevent {\bf Wd2} from witnessing the
uncommitted writes of transaction {\bf Wd1}.  Subsequently committing
{\bf Wd1} leads to the loss of {\bf Wd2}'s updates (the so called
\emph{lost update anomaly}~\cite{berenson}), resulting in an incorrect
balance of \C{k-a1}. The second execution describes a similar scenario
with {\bf Wd1} and {\bf Wd2} exchanging their roles.  Clearly, {\sc
  rc} is an excessively weak isolation level for this program because
it loses the updates of one transaction, resulting in the violation of
the post-condition.  A stronger isolation level that prevents lost
updates is required.  \iso{Snapshot Isolation} ({\sc
  si})~\cite{berenson} fits this requirement; {\sc si} effectively
serializes transactions that update a shared data object by aborting
and re-executing a transaction if write-write conflicts are detected
during its commit.\footnote{{\sc si} is not functionally the same as
  {\sc ser}, however; the differences in the two manifest in the \emph{absence}
  of any write-write conflicts.}   Since {\sc si},
unlike {\sc ser}, does not rely on expensive lock-based
concurrency control, it is also more efficient, making it appropriate
appropriate for both {\bf Wd} transactions.

@@

For the rely-guarantee approach to be useful, it has to thus constrain
allowed interference in accordance with the chosen level of isolation
associated with a transaction. The first key observation we make is
that we can constrain interference by constraining the definition of
the \emph{happens-before} ($\hbZ$) relation between transactions, or
their respective constituents.  To do so, we axiomatize the $\hbZ$
relation to capture the interference characteristics of different
isolation levels as their specifications ($\psi$). For example, a
\iso{Read Committed} specification allows an operation $\eta_1$ of
$T_1$ to happen before an operation $\eta_2$ of $T_2$ only if every
operation in $T_1$ (including its commit) happens before $\eta_2$ (we
let $T_1 \hboar \eta_2$ denote this). The {\sc rc} specification
does not require $T_1$ to execute and commit before all operations of
$T_2$, thus allowing it to interfere in $T_2$. The \iso{Snapshot
  Isolation} specification requires two transactions that write to the
same variable to be related by happens-before.  This effectively
prohibits interference due to actions in $T_1$ being interleaved in
$T_2$ (or, vise versa).  Formally\footnote{Note that specifications
    presented here are solely for illustration. The actual
    specifications described in \S\ref{sec:ansi-isolation} are more nuanced.}:
\begin{smathpar}
\begin{array}{lcl}
\psi_{RC} & \defeq & \forall T_1,T_2,\eta_1,\eta_2,.\; \txn(\eta_1) = T_1 
  \conj \txn(\eta_2) = T_2 \\
  & & \hspace*{0.6in}\conj T_1 \neq T_2 \conj \eta_1 \hboar
  \eta_2 \Rightarrow T_1 \hboar \eta_2 \\
\psi_{SI} & \defeq & \forall T_1,T_2.\; T_1 \neq T_2 \conj
  (\exists \C{X}.~{T_1 \wrstoar \C{X}} \conj 
                      {T_2 \wrstoar \C{X}})\\
  &  & \hspace*{0.6in}\Rightarrow{T_1 \hboar T_2} \disj {T_2 \hboar T_1} \\
\end{array}
\end{smathpar}
Fig.~\ref{fig:motiv-eg-1-hb} is a visualization of $\hbZ$ 
from {\bf Wd2} to {\bf Wd1} allowed by $\psi_{RC}$ and $\psi_{SI}$. All
arrows are legal under $\psi_{RC}$ because in no case does an operation
$\eta_1$ from {\bf Wd1} happen-before $\eta_2$ of {\bf Wd2} without the commit
of {\bf Wd1} also happening before $\eta_2$. $\psi_{SI}$ however disallows
all dashed arrows, because dashed arrows establish $\hbZ$ between actions in {\bf Wd2}
and only a subset of the operations in {\bf Wd1}. Ligher (blue) arrows with
hollow heads denote $\hbZ$ relationships that do not affect the value
of \C{C} in a way that causes the program to violate its
post-condition.  Darker (black) arrows with solid heads denote $\hbZ$
relationships that lead to the violation of the post-condition. The task of
the reasoning framework is to determine if all $\hbZ$ relationships
allowed by an isolation level lead to the satisfaction of the
post-condition.

@@

The specification of an isolation level encodes its interference
characteristic as constraints over the $\hbZ$ relation. However, for this
to be useful in reasoning about programs, the rely-guarantee framework
should be able to use the specification to determine if an
interference is valid or invalid, allowing the programmer to only
focus on the former. Our second key observation is that this is
possible if the reasoning framework adequately tracks $\hbZ$ at each
program point, while preserving $\hbZ$ constraints as invariants
between program points. An interference that leads to the violation of
$\hbZ$ constraints (i.e., an invalid interference) is thus
automatically prohibited. For instance, consider the program point
after the write to \C{C} in {\bf Wd1}. The expected invariant ($\phi$) at
that program point is shown below ($\committed$ stands for
``committed''):
\begin{smathpar}
\begin{array}{l}
  \neg\committed(\C{Wd1}) \conj (\neg\committed(\C{Wd2}) \Rightarrow
  \C{C = k-a1}) 
                \\
      \C{Wd1} \wrstoar \C{C} \conj (\committed(\C{Wd2})
                \Rightarrow \C{C = k-a1-a2})\}
\end{array}
\end{smathpar}
$\phi$ asserts that {\bf Wd1} is not yet committed, and that it wrote to
\C{C}, and the value of \C{C} is either \C{k-a1-a2} or \C{k-a1}
depending on whether or not {\bf Wd2} is committed. If $\phi$ remains
invariant until {\bf Wd1} commits, then the post-condition (\C{C =
k-a1-a2}) can be established easily. However, an interference from
{\bf Wd2} at this stage (captured by the last dashed arrow in
Fig.~\ref{fig:motiv-eg-1-hb}) may violate the invariant by writing
$\C{k-a1}$ to \C{C} and committing {\bf Wd2}, thus leading to
$\committed(\C{Wd2}) \Rightarrow \C{C=k-a2}$. Fortunately,
\iso{Snapshot Isolation} prevents this interference, and this can be
shown by demonstrating that an interference from {\bf Wd2} starting from an
execution state that satisfies $\psi_{SI} \wedge \phi$ leads to an
execution state where neither $\C{Wd2} \hboar \C{Wd1}$ nor $\C{Wd1}
\hboar \C{Wd2}$ holds (the former does not hold because at least one write
from {\bf Wd1} has happened before at least one operation of {\bf Wd2}, and
the latter does not hold because {\bf Wd1} has not yet committed, and {\bf Wd2}
has already begun). Since {\bf Wd2} also writes to \C{C}, this violates
the $\psi_{SI}$ constraint which we assume to be an invariant. A proof
that the post-condition holds now follows from the contradiction. It is
informative to note that if the invariant is $\psi_{RC}$ instead of
$\psi_{SI}$, we cannot derive a contradiction and we cannot rule out
the interference, which (rightfully) causes  the proof to fail.

>>>>

\begin{figure}[!h]
\centering
\subcaptionbox {
  {\sc rc} Execution 1
  \label{fig:motiv-eg-1-a}
} [
  0.55\columnwidth
] {
  \includegraphics[scale=0.5]{Figures/motiv-eg-1-a}
}
%\hspace*{0.5in}
\subcaptionbox {
  {\sc rc} Execution 2
  \label{fig:motiv-eg-1-b}
}{
  \includegraphics[scale=0.5]{Figures/motiv-eg-1-b}
}
\caption{A possible execution of the program shown in Fig.~\ref{fig:motiv-eg-1} under
  a read committed isolation level.}
\label{fig:rc-executions}
\end{figure}
The figure depicts an execution as a series of read, write, and commit
operations.  The subscript of an operation indicates the transaction
that executes it\footnote{For clarity, the effects of different
  transactions are shown in different colored backgrounds.} In the
first execution, transaction {\bf Wd1} reads the current balance
(\C{k}) and writes the new balance (\C{k-a1}), but before it commits
transaction {\bf Wd2} executes and commits, writing the new balance
(\C{k-a2}). RC isolation prevent {\bf Wd2} from witnessing the
uncommitted writes of transaction {\bf Wd1}.  Subsequently committing
{\bf Wd1} leads to the loss of {\bf Wd2}'s updates (the so called
\emph{lost update anomaly}~\cite{berenson}), resulting in an incorrect
balance of \C{k-a1}. The second execution describes a similar scenario
with {\bf Wd1} and {\bf Wd2} exchanging their roles.  Clearly, {\sc
  rc} is an excessively weak isolation level for this program because
it loses the updates of one transaction, resulting in the violation of
the post-condition.  A stronger isolation level that prevents lost
updates is required.  \iso{Snapshot Isolation} ({\sc
  si})~\cite{berenson} fits this requirement; {\sc si} effectively
serializes transactions that update a shared data object by aborting
and re-executing a transaction if write-write conflicts are detected
during its commit.\footnote{{\sc si} is not functionally the same as
  {\sc ser}, however; the differences in the two manifest in the \emph{absence}
  of any write-write conflicts.}   Since {\sc si},
unlike {\sc ser}, does not rely on expensive lock-based
concurrency control, it is also more efficient, making it appropriate
appropriate for both {\bf Wd} transactions.

\begin{figure}
\begin{smathpar}
\begin{array}{lcl}
\psi_{RC} & \defeq & \forall T_1,T_2,\eta_1,\eta_2,.\; \txn(\eta_1) = T_1 
  \conj \txn(\eta_2) = T_2 \\
  & & \hspace*{0.6in}\conj T_1 \neq T_2 \conj \eta_1 \hboar
  \eta_2 \Rightarrow T_1 \hboar \eta_2 \\
\psi_{SI} & \defeq & \forall T_1,T_2.\; T_1 \neq T_2 \conj
  (\exists \C{X}.~{T_1 \wrstoar \C{X}} \conj 
                      {T_2 \wrstoar \C{X}})\\
  &  & \hspace*{0.6in}\Rightarrow{T_1 \hboar T_2} \disj {T_2 \hboar T_1} \\
\end{array}
\end{smathpar}
\caption{Visibility and interference properties for different weak isolation levels can be
  captured by different instantiations of their happens-before relation.}
\label{fig:interference-ex}
\end{figure}

@@

For the rely-guarantee approach to be useful, it has to thus constrain
allowed interference in accordance with the chosen level of isolation
associated with a transaction. The first key observation we make is
that we can constrain interference by constraining the definition of
the \emph{happens-before} ($\hbZ$) relation between transactions, or
their respective constituents.  To do so, we axiomatize the $\hbZ$
relation to capture the interference characteristics of different
isolation levels as their specifications ($\psi$). For example, a
\iso{Read Committed} specification allows an operation $\eta_1$ of
$T_1$ to happen before an operation $\eta_2$ of $T_2$ only if every
operation in $T_1$ (including its commit) happens before $\eta_2$ (we
let $T_1 \hboar \eta_2$ denote this). The {\sc rc} specification
does not require $T_1$ to execute and commit before all operations of
$T_2$, thus allowing it to interfere in $T_2$. The \iso{Snapshot
  Isolation} specification requires two transactions that write to the
same variable to be related by happens-before.  This effectively
prohibits interference due to actions in $T_1$ being interleaved in
$T_2$ (or, vise versa).  We formalize these properties in Fig.~\ref{fig:interference-ex}\footnote{Note that specifications
    presented here are solely for illustration. The actual
    specifications described in \S\ref{sec:ansi-isolation} are more nuanced.}.

Fig.~\ref{fig:motiv-eg-1-hb} is a visualization of $\hbZ$ 
from {\bf Wd2} to {\bf Wd1} allowed by $\psi_{RC}$ and $\psi_{SI}$. All
arrows are legal under $\psi_{RC}$ because in no case does an operation
$\eta_1$ from {\bf Wd1} happen-before $\eta_2$ of {\bf Wd2} without the commit
of {\bf Wd1} also happening before $\eta_2$. $\psi_{SI}$ however disallows
all dashed arrows, because dashed arrows establish $\hbZ$ between actions in {\bf Wd2}
and only a subset of the operations in {\bf Wd1}. Ligher (blue) arrows with
hollow heads denote $\hbZ$ relationships that do not affect the value
of \C{C} in a way that causes the program to violate its
post-condition.  Darker (black) arrows with solid heads denote $\hbZ$
relationships that lead to the violation of the post-condition. The task of
the reasoning framework is to determine if all $\hbZ$ relationships
allowed by an isolation level lead to the satisfaction of the
post-condition.

@@

The specification of an isolation level encodes its interference
characteristic as constraints over the $\hbZ$ relation. However, for this
to be useful in reasoning about programs, the rely-guarantee framework
should be able to use the specification to determine if an
interference is valid or invalid, allowing the programmer to only
focus on the former. Our second key observation is that this is
possible if the reasoning framework adequately tracks $\hbZ$ at each
program point, while preserving $\hbZ$ constraints as invariants
between program points. An interference that leads to the violation of
$\hbZ$ constraints (i.e., an invalid interference) is thus
automatically prohibited. For instance, consider the program point
after the write to \C{C} in {\bf Wd1}. The expected invariant ($\phi$) at
that program point is shown below ($\committed$ stands for
``committed''):

\begin{smathpar}
\begin{array}{l}
  \neg\committed(\C{Wd1}) \conj (\neg\committed(\C{Wd2}) \Rightarrow
  \C{C = k-a1}) 
                \\
      \C{Wd1} \wrstoar \C{C} \conj (\committed(\C{Wd2})
                \Rightarrow \C{C = k-a1-a2})\}
\end{array}
\end{smathpar}

\noindent $\phi$ asserts that {\bf Wd1} is not yet committed, and that it wrote to
\C{C}, and the value of \C{C} is either \C{k-a1-a2} or \C{k-a1}
depending on whether or not {\bf Wd2} is committed. If $\phi$ remains
invariant until {\bf Wd1} commits, then the post-condition (\C{C =
k-a1-a2}) can be established easily. However, an interference from
{\bf Wd2} at this stage (captured by the last dashed arrow in
Fig.~\ref{fig:motiv-eg-1-hb}) may violate the invariant by writing
$\C{k-a1}$ to \C{C} and committing {\bf Wd2}, thus leading to
$\committed(\C{Wd2}) \Rightarrow \C{C=k-a2}$. Fortunately,
\iso{Snapshot Isolation} prevents this interference, and this can be
shown by demonstrating that an interference from {\bf Wd2} starting from an
execution state that satisfies $\psi_{SI} \wedge \phi$ leads to an
execution state where neither $\C{Wd2} \hboar \C{Wd1}$ nor $\C{Wd1}
\hboar \C{Wd2}$ holds (the former does not hold because at least one write
from {\bf Wd1} has happened before at least one operation of {\bf Wd2}, and
the latter does not hold because {\bf Wd1} has not yet committed, and {\bf Wd2}
has already begun). Since {\bf Wd2} also writes to \C{C}, this violates
the $\psi_{SI}$ constraint which we assume to be an invariant. A proof
that the post-condition holds now follows from the contradiction. It is
informative to note that if the invariant is $\psi_{RC}$ instead of
$\psi_{SI}$, we cannot derive a contradiction and we cannot rule out
the interference, which (rightfully) causes  the proof to fail.

### 2016-06-23 opsem.tex
edits

\section{\txnimp}
\label{sec:opsem}

@@

\subsection{Syntax and Semantics}
\label{sec:syntax}

@@

Fundamental to our development is the notion of a trace invariant
($\I$). $\I$ is a function from traces ($\E$) to propositions
($\texttt{Prop}$) that define well-formedness constraints over traces.
The machine takes a step only if the resulting trace satisfies the
constraints imposed by $\I$. This behaviour is captured by the
auxiliary reduction rule \rulelabel{E-Aux} that factors out the trace
extension aspect of the evaluation by abstracting away the
operation-specific behaviour as a function that generates an
appropriate effect. We let $\mathcal{F}$ denote this function.
\rulelabel{E-Aux} uses $\mathcal{F}$ to generate a new effect and
extend the trace ($\E = (\A,\visZ)$) \emph{only if} the
well-formedness constraints imposed by $\I$ on $\E$ (i.e., $\I(\E')$)
are satisfied. Otherwise, it gets stuck. In an execution that runs to
completion, every small-step preserves the well-formedness of a trace,
thus ensuring the invariance of $\I$.  Note that the semantics makes
no assumptions about $\I$ other than its type. As such, it can be
instantiated with any trace-parametric proposition that expresses
constraints over the given trace. For instance, consider the
$\psi_{RC}$ specification from \S\ref{sec:motivation}, but with
bounded $T_1$ and $T_2$ instantiated with \C{Wd1} and \C{Wd2},
respectively. The instantiated specification is the following
proposition:
\begin{smathpar}
\begin{array}{l}
  \forall \eta_1,\eta_2,.\; \txn(\eta_1) = \C{Wd1}
  \conj \txn(\eta_2) = \C{Wd2} \\
  \hspace*{0.6in}\conj \C{Wd1} \neq \C{Wd2} \conj \eta_1 \hboar
  \eta_2 \Rightarrow \C{Wd1} \hboar \eta_2 \\
\end{array}
\end{smathpar}
It is easy to intepret the above proposition in the context of a trace
$\E$ that captures an execution of the program in
Fig.~\ref{fig:motiv-eg-1}. Such a trace-parametric proposition can be
used to instantiate the trace invariant $\I$ in Fig.~\ref{fig:txnimp}.
The resultant operational semantics describes an abstract machine that
gets stuck if an operation of \C{Wd2} is executed in a state that
incorporates some, but not all the effects (including the \C{COMMIT})
of \C{Wd1}. 
% While the machine takes a step only if the constraints are
% satisfied, it neither defines nor explicitly assumes an oracle to
% check satisfaction.

@@

As described previously, \rulelabel{E-Aux} abstracts away the
operation-specific behaviour of a machine step as a function ($\F$)
that accepts a set ($S$) of effects chosen by the machine to make
visible to the operation, interprets the operation w.r.t $S$, and
returns an appropriate effect that encodes its return value. Rules
\rulelabel{E-Var}, \rulelabel{E-Asgn} and \rulelabel{E-Commit} define
such functions for read, write and commit operations, respectively.
The effect returned by the function in each case includes its
transaction id ($T_i$) along with an arbitrarily chosen effect id
($j$) that is later verified to be unique in \rulelabel{E-Aux}. The
$\rval$ for a write is the value being written, and for commit it is
$\bot$. In case of a read, the value read depends on the how the read
operation chooses to interpret the given set ($S$) of visible effects.
The interpretation may depend on the application semantics. For
example, a monotonically increasing counter application may choose to
let a write with the largest value determine the value of a
read. A more straightforward interpretation has already been described
before. To accommodate multiple interpretations, the semantics
is made parametric over an interpretation function ($\interp{\cdot}$)
that accepts a set of effects and a variable name, and returns the
value associated with the variable. A straightforward interpretation
function that chooses the last write (i.e., write with largest id) is
shown below:
\begin{smathpar}
\begin{array}{lcl}
  \isMax(S,\eta) & \Leftrightarrow &  \forall (\eta'\in S).  
  \kind(\eta') = \kind(\eta) \\
  & & \hspace*{0.4in}\Rightarrow \eta' = \eta \disj \id(\eta') < \id(\eta)\\

\interp{S}(X) & = & \C{if}\;(\exists (\eta \in S). \kind(\eta) = \C{WR}(X) 
  \wedge \isMax(S,\eta)) \\
  & & \C{then}\;\rval(\eta)\;\C{else}\;0\\
\end{array}
\end{smathpar}
Rules \rulelabel{E-Top-Ctx} and \rulelabel{E-Txn-Ctx} define
congruence properties for top-level terms and transaction-bound terms,
respectively. The rules and evaluation contexts ($\ectx$ and
$\ectx_i$) are defined such that only certain kinds of terms are
allowed at the top-level and inside a transaction. In particular, a
\txnimp program at the top-level can either be a transaction, or a
parallel composition of transactions. A command inside a \C{txn} block
can either be an assignment, or a sequential composition of
assignments.

>>>>

\subsection{\txnimp: Syntax and Semantics}
\section{\txnimp}
\label{sec:opsem}

\label{sec:syntax}

@@

@@

Fundamental to our development is the notion of a trace invariant
($\I$). $\I$ is a function from traces ($\E$) to propositions
($\texttt{Prop}$) that define well-formedness constraints over traces.
The machine takes a step only if the resulting trace satisfies the
constraints imposed by $\I$. This behaviour is captured by the
auxiliary reduction rule \rulelabel{E-Aux} that factors out the trace
extension aspect of the evaluation by abstracting away the
operation-specific behaviour as a function that generates an
appropriate effect. We let $\mathcal{F}$ denote this function.
\rulelabel{E-Aux} uses $\mathcal{F}$ to generate a new effect and
extend the trace ($\E = (\A,\visZ)$) \emph{only if} the
well-formedness constraints imposed by $\I$ on $\E$ (i.e., $\I(\E')$)
are satisfied. Otherwise, it gets stuck. In an execution that runs to
completion, every small-step preserves the well-formedness of a trace,
thus ensuring the invariance of $\I$.  Note that the semantics makes
no assumptions about $\I$ other than its type. As such, it can be
instantiated with any trace-parametric proposition that expresses
constraints over the given trace. For instance, consider the
$\psi_{RC}$ specification from \S\ref{sec:motivation}, but with
bounded $T_1$ and $T_2$ instantiated with \C{Wd1} and \C{Wd2},
respectively. The instantiated specification is the following
proposition:

\begin{smathpar}
\begin{array}{l}
  \forall \eta_1,\eta_2,.\; \txn(\eta_1) = \C{Wd1}
  \conj \txn(\eta_2) = \C{Wd2} \\
  \hspace*{0.6in}\conj \C{Wd1} \neq \C{Wd2} \conj \eta_1 \hboar
  \eta_2 \Rightarrow \C{Wd1} \hboar \eta_2 \\
\end{array}
\end{smathpar}

\noindent It is easy to intepret the above proposition in the context of a trace
$\E$ that captures an execution of the program in
Fig.~\ref{fig:motiv-eg-1}. Such a trace-parametric proposition can be
used to instantiate the trace invariant $\I$ in Fig.~\ref{fig:txnimp}.
The resultant operational semantics describes an abstract machine that
gets stuck if an operation of \C{Wd2} is executed in a state that
incorporates some, but not all the effects (including the \C{COMMIT})
of \C{Wd1}. 
% While the machine takes a step only if the constraints are
% satisfied, it neither defines nor explicitly assumes an oracle to
% check satisfaction.

@@

As described previously, \rulelabel{E-Aux} abstracts away the
operation-specific behaviour of a machine step as a function ($\F$)
that accepts a set ($S$) of effects chosen by the machine to make
visible to the operation, interprets the operation w.r.t $S$, and
returns an appropriate effect that encodes its return value. Rules
\rulelabel{E-Var}, \rulelabel{E-Asgn} and \rulelabel{E-Commit} define
such functions for read, write and commit operations, respectively.
The effect returned by the function in each case includes its
transaction id ($T_i$) along with an arbitrarily chosen effect id
($j$) that is later verified to be unique in \rulelabel{E-Aux}. The
$\rval$ for a write is the value being written, and for commit it is
$\bot$. In case of a read, the value read depends on the how the read
operation chooses to interpret the given set ($S$) of visible effects.
The interpretation may depend on the application semantics. For
example, a monotonically increasing counter application may choose to
let a write with the largest value determine the value of a
read. A more straightforward interpretation has already been described
before. To accommodate multiple interpretations, the semantics
is made parametric over an interpretation function ($\interp{\cdot}$)
that accepts a set of effects and a variable name, and returns the
value associated with the variable. A straightforward interpretation
function that chooses the last write (i.e., write with largest id) is
shown below:

\begin{smathpar}
\begin{array}{lcl}
  \isMax(S,\eta) & \Leftrightarrow &  \forall (\eta'\in S).  
  \kind(\eta') = \kind(\eta) \\
  & & \hspace*{0.4in}\Rightarrow \eta' = \eta \disj \id(\eta') < \id(\eta)\\

\interp{S}(X) & = & \C{if}\;(\exists (\eta \in S). \kind(\eta) = \C{WR}(X) 
  \wedge \isMax(S,\eta)) \\
  & & \C{then}\;\rval(\eta)\;\C{else}\;0\\
\end{array}
\end{smathpar}

\noindent Rules \rulelabel{E-Top-Ctx} and \rulelabel{E-Txn-Ctx} define
congruence properties for top-level terms and transaction-bound terms,
respectively. The rules and evaluation contexts ($\ectx$ and
$\ectx_i$) are defined such that only certain kinds of terms are
allowed at the top-level and inside a transaction. In particular, a
\txnimp program at the top-level can either be a transaction, or a
parallel composition of transactions. A command inside a \C{txn} block
can either be an assignment, or a sequential composition of
assignments.
